---
title: "Suicide rates in the world from 1985 to 2016"
author: "Phuong K. Vu"
date: "12/06/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(ggplot2)
setwd("~/Desktop/UW/Autumn 2019/INFO 201/extra-credit")
source('analysis.R')
```

## Introduction
The goal of this project is to illustrate the suicide situation in the world from 1985 to 2016. Hopefully from the data, graph and insight of the graph, the society and government would come up with plans and solutions to prevent the growth of suicides rate and ensure better psychological life of the pepople. The dataset for this project was collected over a long period of time (1985 - 2016) from different countries in the world. The final dataset was pulled from four other datasets, which were cited in the reference section and conducted by United Nation Development Program, World Bank, Szamil and World Health Organization. 

## Data Graphic
Below is the time series plot of the original data. Each point in the graph represent the total of suicide cases in the world in each specific year and categorized by gender. 

```{r echo= FALSE, message= FALSE}
draw_plot()
```

## Discussion 
From the time series plot above, it is obvious that men are much more likely to commit suicide than women. In addition, the number of suicide case has a trend to slowly increase over time. However, there was a dramatical drop in the number in 2016, which states that people's psychological health has initial improvement that made them to continue living. The local government and community leaders are definitely the indirect stakeholders of this project, because they are urged to create some changes to help the community solve current suicidal problem. Meanwhile, people who are suffering psychological diseases and losing their desire to live are the direct stakeholders because this project may result in significant changes in social health system that would provide more support to help them get better from their mental illness. 

## Reflection
From the analysis of the time series plot, I learned that psychological disorders might be disregarded and ignored by the society before 21st century. Therefore, there were a significant increase in sucide cases every year before 2000. However, as people became more opened with multipple aspects of the community, the treatment toward mental-illed people was getting better and there were more psychiatrist to help cure them. As a result, I learned that how the community embrace and accept psychological disorders would have a significant impact upon the number of suicide cases in that community. Indeed, after this project, there is another question raised in my mind about what might be other factors that would influence the suicide situation in a community. In the future, I would love to conduct other projects that explore this field to find the answer to my quesiton. 

## Code Appendix
In order to manipulate the data and create the plot for this project, I created the following R code files: <br>
- analysis.R : code file of function to execute tasks <br>
- master.csv : original dataset <br>
- README.md : introduction file with affiliation <br>
- Report.Rmd : the actual report file <br>
Link to my Github repository of this project: [Github repository](https://github.com/pvu-1864424/extra-credit.git)

## Reference
The compiled dataset is pulled from four other datasets linked by time and place, and was built to find signals correlated to increased suicide rates among different cohorts globally, across the socio-economic spectrum: <br>
United Nations Development Program. (2018). Human development index (HDI). Retrieved from http://hdr.undp.org/en/indicators/137506 <br>
World Bank. (2018). World development indicators: GDP (current US$) by country:1985 to 2016. Retrieved from http://databank.worldbank.org/data/source/world-development-indicators# <br>
[Szamil]. (2017). Suicide in the Twenty-First Century [dataset]. Retrieved from https://www.kaggle.com/szamil/suicide-in-the-twenty-first-century/notebook <br>
World Health Organization. (2018). Suicide prevention. Retrieved from http://www.who.int/mental_health/suicide-prevention/en/ <br>

